---
title: "Regression Models Wee-3"
author: "Saul Lugo"
date: "December 25, 2015"
output: html_document
---

#Multivariable Regression

##Loading the data
```{r loading_data}
require(datasets)
data(swiss)
require(GGally) #this library provides some additional plots types to ggplot2
require(ggplot2)

g = ggpairs(swiss, lower = list(continuous = "smooth"), params = c(method = "looess"))
g
```

##Fit all of the variable in the same Linear Model

```{r some_lm}
summary(lm(Fertility ~ ., data = swiss))$coefficients
```

##Interpretation of the coefficients
For example, look at the coefficient for Agriculture. It is -0.17. We can interpret it as: **our model estimates an expected 0.17 decrease in standarized fertility for every 1% increase in percentage of males involved in agriculture in holding the remaining variables constant.***

##What happen if we take only one of the variables as a predictor of the Ferlity
For example, let's fit the model taking only the variable Agriculture as the regressor:

```{r fit_one_variable}
summary(lm(Fertility ~ Agriculture, data = swiss))$coefficients
```

Now Agriculture has a positive correlation with fertility. This conclusion might be wrong.

